Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2019-01-29 WholePlotTrajectories.Rmd
intro

The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, the carbon, water and nutrient cycles, and ensure cultural and human well-being. 
The growing demand in forests products together with current global changes, however, increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
This context specifically threatens the maintenance of community structure, composition and functioning [@Anderson-Teixeira2013].
In tropical forests indeed, tree communities biodiversity and composition rely on a natural disturbance regime, specifically through tree fall gaps [@Schnitzer2001].
This disturbance regime is affected by the current changing context and anthropic activities [@Sist2015], so clarify community response to disturbance and the underlying ecological processes is crucial to anticipate the fate of tropical forests [@Chazdon2003].

Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and modifications in species interactions, mainly through the reduction of species density [@Chesson2000]. 
For now, community response to disturbance has been largely studied through forest structural parameters, rapid and convenient to measure, such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
Thereafter, models based on the observed trajectories after disturbance gave important insights into community recovery and resilience [@Herault2018]. 
Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood.
A leading idea on how disturbance affect community diversity is the Intermediate Disturbance Hypothesis (IDH) [@Connell1978] predicting biodiversity peaks at intermediate value of disturbance intensity and at intermediate time after disturbance. 
The IDH is founded on the emergence after disturbance of deterministic processes that prevent competitive exclusion through the fluctuations of dominant competitors along time due to disturbance environmental impacts [@Fox2013].
The IDH assumes predictable post-disturbance trajectories depending on the intensity of disturbance, that define the importance of disturbance impacts [@Willig2018].
It also assumes, without additional disturbance, a recovery of the pre-disturbance state defined by the dominance of late-successional species [@Rees2001;@Carreno-Rocabado2012;@Poorter2016].
In highly diverse tropical forests, however, empirical validation of the IDH remains debated [@Fox2013;@Sheil2013] and observed post-disturbance trajectories are often blurry compared to theoretical expectations [@Hugues2007].

>>>>

The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, the carbon, water and nutrient cycles, and ensure cultural and human well-being. 
The growing demand in forests products together with current global changes, however, increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
This changing context alters the natural disturbance regime that shape and maintain tree community structure, composition and functioning [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015], so anticipate the fate of tropical forests requires to understand community response to disturbance.
Disturbance initiate ecological succession comprising a suit of changes in community structure and functioning over time, which rate, endpoint an duration depend on the disturbance intensity [@Chazdon2003;@Willig2018].
Regarding tropical forests, community response to disturbance would be secondary succession, starting with an intact soil and a consistent seed bank, from which we aim would be to highlight predictable trajectories and their underlying ecological processes.

Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and modifications in species interactions, mainly through modifications of species density [@Chesson2000]. 
For now, post-disturbance response has been largely studied through forest structural parameters, rapid and convenient to measure, such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
Thereafter, models based on the observed trajectories after disturbance gave important insights into community recovery and resilience [@Herault2018]. 
Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood.
A leading idea on how the impacts of disturbance on community diversity is the Intermediate Disturbance Hypothesis (IDH) [@Connell1978].
The IDH assumes biodiversity changes in relation to disturbance intensity and frequency, with diversity peaks for intermediate intensity and at intermediate time after disturbance. 
At intermediate level of disturbance, the fluctuation of environmental conditions entails deterministic processes favoring both competitively superior species and fast colonizers, hence preventing competitive exclusion [@Shea2004;@Pulsford2016].

The IDH then assumes both predictable trajectories depending on the disturbance intensity, and, without additional disturbance, the recovery of pre-disturbance state defined by the dominance of late-successional species [@Rees2001;@Sheil2003;@Willig2018].
In highly diverse tropical forests, however, empirical validation of the IDH remains debated [@Fox2013;@Sheil2013] and observed post-disturbance trajectories are often blurry compared to theoretical expectations [@Hugues2007].

### 2019-01-29 WholePlotTrajectories.tex
intro

The large areas covered with tropical forests worldwide hold crucial
environmental, economic and social values. They provide wood and
multiple non-timber forest products, shelter a diversified fauna,
regulate the local and regional climates, the carbon, water and nutrient
cycles, and ensure cultural and human well-being. The growing demand in
forests products together with current global changes, however, increase
the pressure on remaining natural forests \citep{Morales-Hidalgo2015}.
This context specifically threatens the maintenance of community
structure, composition and functioning \citep{Anderson-Teixeira2013}. In
tropical forests indeed, tree communities biodiversity and composition
rely on a natural disturbance regime, specifically through tree fall
gaps \citep{Schnitzer2001}. This disturbance regime is affected by the
current changing context and anthropic activities \citep{Sist2015}, so
clarify community response to disturbance and the underlying ecological
processes is crucial to anticipate the fate of tropical forests
\citep{Chazdon2003}.

Disturbance change both community abiotic and biotic environments,
through modifications in the fluxes of light, heat and water
\citep{Goulamoussene2017}, and modifications in species interactions,
mainly through the reduction of species density \citep{Chesson2000}. For
now, community response to disturbance has been largely studied through
forest structural parameters, rapid and convenient to measure, such as
aboveground biomass, tree height or stem density
\citep{Piponiot2016, Rutishauser2016}. Thereafter, models based on the
observed trajectories after disturbance gave important insights into
community recovery and resilience \citep{Herault2018}. Regarding
community diversity and composition, however, post-disturbance
trajectories have not been as thoroughly understood. A leading idea on
how disturbance affect community diversity is the Intermediate
Disturbance Hypothesis (IDH) \citep{Connell1978} predicting biodiversity
peaks at intermediate value of disturbance intensity and at intermediate
time after disturbance. The IDH is founded on the emergence after
disturbance of deterministic processes that prevent competitive
exclusion through the fluctuations of dominant competitors along time
due to disturbance environmental impacts \citep{Fox2013}. The IDH
assumes predictable post-disturbance trajectories depending on the
intensity of disturbance, that define the importance of disturbance
impacts \citep{Willig2018}. It also assumes, without additional
disturbance, a recovery of the pre-disturbance state defined by the
dominance of late-successional species
\citep{Rees2001, Carreno-Rocabado2012, Poorter2016}. In highly diverse
tropical forests, however, empirical validation of the IDH remains
debated \citep{Fox2013, Sheil2013} and observed post-disturbance
trajectories are often blurry compared to theoretical expectations
\citep{Hugues2007}.

>>>>

The large areas covered with tropical forests worldwide hold crucial
environmental, economic and social values. They provide wood and
multiple non-timber forest products, shelter a diversified fauna,
regulate the local and regional climates, the carbon, water and nutrient
cycles, and ensure cultural and human well-being. The growing demand in
forests products together with current global changes, however, increase
the pressure on remaining natural forests \citep{Morales-Hidalgo2015}.
This changing context alters the natural disturbance regime that shape
and maintain tree community structure, composition and functioning
\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}, so anticipate
the fate of tropical forests requires to understand community response
to disturbance. Disturbance initiate ecological succession comprising a
suit of changes in community structure and functioning over time, which
rate, endpoint an duration depend on the disturbance intensity
\citep{Chazdon2003, Willig2018}. Regarding tropical forests, community
response to disturbance would be secondary succession, starting with an
intact soil and a consistent seed bank, from which we aim would be to
highlight predictable trajectories and their underlying ecological
processes.

Disturbance change both community abiotic and biotic environments,
through modifications in the fluxes of light, heat and water
\citep{Goulamoussene2017}, and modifications in species interactions,
mainly through modifications of species density \citep{Chesson2000}. For
now, post-disturbance response has been largely studied through forest
structural parameters, rapid and convenient to measure, such as
aboveground biomass, tree height or stem density
\citep{Piponiot2016, Rutishauser2016}. Thereafter, models based on the
observed trajectories after disturbance gave important insights into
community recovery and resilience \citep{Herault2018}. Regarding
community diversity and composition, however, post-disturbance
trajectories have not been as thoroughly understood. A leading idea on
how the impacts of disturbance on community diversity is the
Intermediate Disturbance Hypothesis (IDH) \citep{Connell1978}. The IDH
assumes biodiversity changes in relation to disturbance intensity and
frequency, with diversity peaks for intermediate intensity and at
intermediate time after disturbance. At intermediate level of
disturbance, the fluctuation of environmental conditions entails
deterministic processes favoring both competitively superior species and
fast colonizers, hence preventing competitive exclusion
\citep{Shea2004, Pulsford2016}.

The IDH then assumes both predictable trajectories depending on the
disturbance intensity, and, without additional disturbance, the recovery
of pre-disturbance state defined by the dominance of late-successional
species \citep{Rees2001, Sheil2003, Willig2018}. In highly diverse
tropical forests, however, empirical validation of the IDH remains
debated \citep{Fox2013, Sheil2013} and observed post-disturbance
trajectories are often blurry compared to theoretical expectations
\citep{Hugues2007}.
